Proof of concept Zstandard user-provided dictionary compression for FASTA files

### 2022-11-24 paper/paper.qmd
Add CodecLz4

---
title: "Proof of concept Zstandard custom compression dictionary for FASTA files"
format:
  plos-pdf:
    number-sections: false
    journal:
      # This is an identifier for the target journal: 
      # from https://plos.org/resources/writing-center/ following submission guidelines link, the identifier is the part of the URL after https://journals.plos.org/<id>/s/submission-guidelines
      id: plosbiology
    include-in-header: |
      % Remove comment for double spacing
      % \usepackage{setspace} 
      % \doublespacing
author:
  - name: Michael Persico
    equal-contributor: false
    affiliations:
      - ref: aff1
    notes: |
      "Current Address: Dept/Program/Center, Institution Name, City, State, Country"
    corresponding: false
    email: correspondingauthor@institute.edu
author-notes:
  equal-contributor: These authors contributed equally to this work.
  deceased: Deceased
  group: Membership list can be found in the Acknowledgments sections
affiliations:
  - id: aff1
    name: Department of Biology, Concordia University
    city: Montreal
    state: Quebec
    country: Canada
abstract: |
  ## Background

  Zstandard (Zstd) represents a lossless data compression algorithm that is highly configurable and is aimed at coupling high compression ratios with fast compression/decompression performance. Previous studies paired specific Zstd configurations with various file formats in bioinformatics to reduce total data volume. This paper presents a "training mode" pipeline, written in the Julia programming language, wherein a custom compression dictionary is generated from a sample FASTA set in order to explore further compression improvements.

@@

@@

I would like to express my sincere gratitude to Professor David Walsh for his advice that helped shape the research as it progressed; My friends and family that supported me throughout my studies; The Julia community for their support before and throughout the writing of this paper; The Quarto developers for producing the Quarto publishing system along with the PLOS template used in this paper; Luca Di Maio and contributors to the Distrobox tool for the ease of setting up the containerized development environments; Maximiliano Sandoval and contributors to the Citations app  for managing the paper's bibliography, and all other persons that had indirectly assisted through their programs and research.

>>>>

---
title: "Proof of concept Zstandard user-provided dictionary compression for FASTA files"
format:
  plos-pdf:
    number-sections: false
    journal:
      # This is an identifier for the target journal: 
      # from https://plos.org/resources/writing-center/ following submission guidelines link, the identifier is the part of the URL after https://journals.plos.org/<id>/s/submission-guidelines
      id: plosbiology
    include-in-header: |
      % Remove comment for double spacing
      % \usepackage{setspace} 
      % \doublespacing
author:
  - name: Michael Persico
    equal-contributor: false
    affiliations:
      - ref: aff1
    notes: |
      "Current Address: Dept/Program/Center, Institution Name, City, State, Country"
    corresponding: false
    email: correspondingauthor@institute.edu
author-notes:
  equal-contributor: These authors contributed equally to this work.
  deceased: Deceased
  group: Membership list can be found in the Acknowledgments sections
affiliations:
  - id: aff1
    name: Department of Biology, Concordia University
    city: Montreal
    state: Quebec
    country: Canada
abstract: |
  ## Background

  Zstandard (Zstd) represents a lossless data compression mechanism that is highly configurable and is aimed at coupling high compression ratios with fast compression/decompression performance. Previous studies have paired specific Zstd configurations with various file formats in bioinformatics to reduce total data volume. This paper presents a "training mode" pipeline, written in the Julia programming language, wherein a custom compression dictionary is generated from a sample FASTA set in order to explore further compression improvements and compare them to the compression performance of Xz, Zlib, Bzip2, and Lz4 compressors.

@@

A list of direct and indirect dependencies can be found in the repository's Manifest.toml file.

@@

I would like to express my sincere gratitude to Professor David Walsh for his advice that helped shape the research as it progressed; My friends and family that supported me throughout my studies; The Julia community for their support before and throughout the writing of this paper; The Quarto developers for producing the Quarto publishing system[@Allaire_Quarto_2022] along with the PLOS template used for this paper; Luca Di Maio and contributors to the Distrobox tool for the ease of setting up the containerized development environments[@maio]; Maximiliano Sandoval and contributors to the Citations app for managing the paper's bibliography[@sandoval], and all other persons that had indirectly assisted through their programs and research.

### 2022-11-24 paper/paper.tex
Add CodecLz4

% Header and Footer with logo
\usepackage{lastpage,fancyhdr}
\usepackage{epstopdf}
%\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[numbers,square,comma]{natbib}
\bibliographystyle{plos2015}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Proof of concept Zstandard custom compression dictionary for FASTA files},
  pdfauthor={Michael Persico},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Proof of concept Zstandard custom compression dictionary
for FASTA
files} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
\\
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
Michael Persico\textsuperscript{1}
\\
\bigskip
\textbf{1} Department of Biology, Concordia
University, Montreal, Quebec, Canada, 
\bigskip

@@

Zstandard (Zstd) represents a lossless data compression algorithm that
is highly configurable and is aimed at coupling high compression ratios
with fast compression/decompression performance. Previous studies paired
specific Zstd configurations with various file formats in bioinformatics
to reduce total data volume. This paper presents a ``training mode''
pipeline, written in the Julia programming language, wherein a custom
compression dictionary is generated from a sample FASTA set in order to
explore further compression improvements.

@@

I would like to express my sincere gratitude to Professor David Walsh
for his advice that helped shape the research as it progressed; My
friends and family that supported me throughout my studies; The Julia
community for their support before and throughout the writing of this
paper; The Quarto developers for producing the Quarto publishing system
along with the PLOS template used in this paper; Luca Di Maio and
contributors to the Distrobox tool for the ease of setting up the
containerized development environments; Maximiliano Sandoval and
contributors to the Citations app for managing the paper's bibliography,
and all other persons that had indirectly assisted through their
programs and research.

>>>>

% Header and Footer with logo
\usepackage{lastpage,fancyhdr}
\usepackage{epstopdf}
%\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[numbers,square,comma]{natbib}
\bibliographystyle{plos2015}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Proof of concept Zstandard user-provided dictionary compression for FASTA files},
  pdfauthor={Michael Persico},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Proof of concept Zstandard user-provided dictionary
compression for FASTA
files} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
\\
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
Michael Persico\textsuperscript{1}
\\
\bigskip
\textbf{1} Department of Biology, Concordia
University, Montreal, Quebec, Canada, 
\bigskip

@@

Zstandard (Zstd) represents a lossless data compression mechanism that
is highly configurable and is aimed at coupling high compression ratios
with fast compression/decompression performance. Previous studies have
paired specific Zstd configurations with various file formats in
bioinformatics to reduce total data volume. This paper presents a
``training mode'' pipeline, written in the Julia programming language,
wherein a custom compression dictionary is generated from a sample FASTA
set in order to explore further compression improvements and compare
them to the compression performance of Xz, Zlib, Bzip2, and Lz4
compressors.

@@

I would like to express my sincere gratitude to Professor David Walsh
for his advice that helped shape the research as it progressed; My
friends and family that supported me throughout my studies; The Julia
community for their support before and throughout the writing of this
paper; The Quarto developers for producing the Quarto publishing
system\citep{Allaire_Quarto_2022} along with the PLOS template used for
this paper; Luca Di Maio and contributors to the Distrobox tool for the
ease of setting up the containerized development
environments\citep{maio}; Maximiliano Sandoval and contributors to the
Citations app for managing the paper's bibliography\citep{sandoval}, and
all other persons that had indirectly assisted through their programs
and research.
